---
title: "A3: SPL Library Checkouts"
name: "Lara Tran"
output: html_document
---

```{r echo = FALSE, message = FALSE, warning = FALSE, code = readLines('summary.R')}
```

### Introduction

In this assignment, I will be analyzing a data set from the Seattle Public Library. The data set the Seattle Library published is based on book titles, authors, checkout year / month, and others collected from the library. I will be answering questions I have about the data set, such as how many books were checked out a year, the most people checked out for a physical book, etc. I want to learn more about the trends of the Seattle public Library book checkout trends which I am hoping to find answers with the values I use.

### Summary Information

One of the 5 values I wanted to find were the most physical book checkouts in the whole data set and that came out to be `r book_checkouts`. Another value I decided to look at is the E-book checkouts from the whole data set and the value I got is `r ebook_checkout`. I also wanted to find the year with the most checkouts which `r checkout_max`. The next value I decided to look for was the maximum checkouts for 2022 which came out to be `r max_checkout`. I also decided to find the year with the lowest checkouts and that is `r min_checkouts`.

### The Dataset

-   Who collected/published the data?

The Seattle Public Library

-   What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?

The parameters of the data are usage class, checkout type, material type, checkout year, checkout month, checkouts, title, ISBN, creator, subjects, publisher, and publication year. In this data set there are 12 columns with 816,354 rows.

-   How was the data collected or generated?

The data is collected monthly by the Seattle Public Library and the data set began on April 2005.

-   Why was the data collected?

The data was collected to see how many books people checked out during the years and what kind of books were collected.

-   What, if any, ethical questions do you need to consider when working with this data?

An ethical question that can be considered when working with this data is whether people that the data was collected from gave consent or knew that the data was being collected.

-   What are possible limitations or problems with this data? (at least 200 words)

A possible limitation or problem with this data set is that this data set is only looking at trends from 2017-2023. If someone were wanting to look at book trends or checkouts from previous years, they wouldn't be able to find it in this data set. The data set was slowly collected from 2017 until now and looks at items checked out at least 10 times a month. This data was only collected from the Seattle Public Library not all libraries in a state or nation. This could mean that if someone wanted to find trends for another state, they wouldn't be able to find them in this data set. One problem I found with this data set was why it was collected. The Seattle Public Data Program only specified why they collect data to improve the public understanding of public operations. I didn't find out why the library data set was collected from the Seattle Public Data Program, but I found it on the Seattle Public Library website stating that the purpose of collecting the data is to reflect on current trends. Overall, I think this data set has many limitations and problems because, like all data sets, not all of them are perfect to fit our needs.

### First Trends Over Time Chart

The first chart I chose is looking at the comparison of ebooks and books from 2017-2023. I think this graph comparison is important because libraries give us a variety of materials we can check out from. In my opinion a physical book and ebook, might be the two that are most checked out. For example, with ebooks people can have it through any technological device. From my personal experience, I haven't checked out a physical book from the library since middle or high school and now if I want to read a book I can read it online with an ebook. In 2018 had the most checkout for books and the lowest in 2023, while in 2020 was the peak for ebook checkouts and in 2023 with the lowest checkouts.

```{r, echo = FALSE, warning = FALSE, code = readLines("chart1.R", message = FALSE)}
```

### Second Trends Over Time Chart

The second chart I chose is a line graph showing the comparison of fiction and nonfiction book checkouts from 2017-2023. Non-fiction compares real world experiences, while fiction is based imaginations. I think this graph is important to look at because fiction books tell an imaginative story and authors create a story that leaves people thinking about hypothetical questions or even cliff hangers. While non-fiction gives us factual information. Looking at the chart, the peak checkout of nonfiction was in 2022 and the lowest in 2017, while the peak for fiction was 2020 and the lowest in 2023.

```{r, echo = FALSE, warning = FALSE, code = readLines("chart2.R"), message = FALSE}
```

### Your Choice: Third Trend Chart

The third chart I chose to do is a line chart. This chart looks at the audiobook checkouts over the time of 2017-2023. I think audiobooks are one important material checkout people can use because you can listen to books anywhere, like on the go, in the car, working out, etc. Audiobooks help people if they have a hard time reading and listening to books is way to comprehend them better. Looking at the chart, the peak checkout of audiobooks was in 2022 and the lowest is in 2023.

```{r echo = FALSE, message = TRUE, warning = FALSE, code = readLines("chart3.R")}
```
